### 2025-01-17 final_project.qmd
update yaml

---
title: "MA Thesis"
author: "Aspen Han"
date: today
format:
  pdf:
    fig-width: 6
    fig-asp: 0.618
    fig-align: center

editor: source

execute:
  echo: false

knitr:
  opts_chunk:
    out-width: "75%"

bibliography: bibliography.bib
csl: <insert desired citation style>

---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
#| label: setup
#| include: false

```

You can add options to executable code like this

```{r}
#| label: raw data
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).